Supporting Information: Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling.

### 2020-05-07 Ben_Manuscripts/stochastic_transport/Supporting_Information.tex
Filling in missing parts of draft

  % mimic this
  \texttt{/setup/parameterize.py}            & 2.1 & Parameterize liquid crystal monomers and solutes with GAFF \\ \hline
  \texttt{/setup/build.py}                   & 2.1 & Build simulation unit cell \\ \hline
  \texttt{/setup/place\_solutes\_pores.py}   & 2.1 & Place equispaced solutes in the pore centers of a unit cell \\\hline
  \texttt{/setup/equil.py}                   & 2.1 & Equilibrate unit cell and run production simulation \\\hline
  \texttt{/analysis/solute\_partitioning.py} & 2.1 & Determine time evolution of partition of solutes between pores and tails \\\hline
  \texttt{/timeseries/msd.py}                & 2.2 & Calculate the mean squared displacement of solutes \\\hline
  \texttt{/analysis/sfbm\_parameters.py}     & 2.2 & Get subordinated fractional Brownian motion parameters by fitting to a solute's dwell and hop length distributions and positional autocorrelation function. \\\hline
  \texttt{/timeseries/ctrwsim.py}            & 2.2 & Generate realizations of a continuous time random walk with the user's choice of dwell and hop distributions \\\hline
  \texttt{/timeseries/forecast\_ctrw.py}     & 2.2 & Combines classes from \texttt{sfbm\_parameters.py} and \texttt{ctrwsim.py} to parameterize and predict MSD in one shot. \\\hline
  \texttt{/analysis/ Markov\_state\_dependent\_dynamics.py} & 2.3 & Identify frame-by-frame state of each solute, construct a transition matrix and simulate realizations of the MSDDM model. \\\hline
  \texttt{/timeseries/mfpt\_pore.py}          & 2.4 & Simulate mean first passage time distributions using the AD approach. \\\hline 
  \end{tabular}

@@

  In Figure~\ref{fig:hurst_autocorrelation}, we plotted Equation~\ref{eqn:fbm_autocorrelation}
  for different values of $H$. When $H > 0.5$, Equation~\ref{eqn:fbm_autocorrelation} decays
  slowly to zero meaning one needs to study large time lags with high frequency in order to
  accurately estimate $H$ from the data. Fortunately, all of our solutes show anti-correlated
  motion, so most of the information in Equation \ref{eqn:fbm_autocorrelation} is contained
  within the first few lags.

@@

  \textit{Determining where to truncate the hop distribution:} A pure
  L\'evy stable distribution has heavy tails which can lead to arbitrarily
  long hop lengths. Our distribution of hop lengths fits well to a L\'evy
  distribution near the mean, but under-samples the tails. In 
  Figure~\ref{fig:truncation_correction} we compare the empirically 
  measured transition emission distribution of the MSDDM for urea to its maximum likelihood fit to
  a L\'evy stable distribution. The ratio between the two distributions 
  at each bin is nearly 1 close to the center, indicating a near-perfect
  fit, larger than 1 slightly further from the center, suggesting that 
  we slightly over sample intermediate hop lengths, and below 1 far from 
  the center, indicating undersampling of extremely long hop lengths.
  Based on the plot, we chose a cut-off of 1 nm in order to compensate for
  over sampled intermediate hop lengths. We chose the same cut-off for all
  solutes.

@@

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{truncation_correction.pdf}
  \caption{We can accurately truncate the marginal distribution of FLM innovations by
  applying a correction to the input truncation parameter. We generated FLM sequences
  and truncated the initial L\'evy stable distribution (before Fourier transforming) at 
  a value of 0.5. After correlation structure is added, the width of the distribution 
  of fractional L\'evy noise decreases significantly. We corrected the input truncation
  parameter with our database resulting in a distribution close to the theoretical
  distribution (black dashed line) with a maximum value close to 0.5.}\label{fig:truncation_correction}
  \end{figure}

@@

  % /supporting_figures/demonstrate_hurst_correction.py
  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{hurst_correction.pdf}
  \caption{Correcting the Hurst parameter input to the algorithm of Stoev and Taqqu
  results in an FLM process with a more accurate correlation structure. We generated
  sequences with an input $H$ of 0.35. We estimated $H$ by fitting the autocorrelation
  function. Without the correction, $H$ is underestimated, meaning realizations are 
  more negatively correlated than they should be.}\label{fig:hurst_correction}
  \end{figure}

@@

  Figure~\ref{fig:anticorrelated_hops} demonstrates that all solutes exhibit the
  same kind of anti-correlated hopping and trapping behavior.

  \begin{figure}[h]
  \centering
  % /figures/distributions_sfbm.py
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_URE.pdf}
  \caption{}\label{fig:hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{URE_powerlaw.pdf}
  \caption{}\label{fig:powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{URE_hop_acf.pdf}
  \caption{}\label{fig:hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_GCL.pdf}
  \caption{}\label{fig:GCL_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{GCL_powerlaw.pdf}
  \caption{}\label{fig:GCL_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{GCL_hop_acf.pdf}
  \caption{}\label{fig:GCL_hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_ACH.pdf}
  \caption{}\label{fig:ACH_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{ACH_powerlaw.pdf}
  \caption{}\label{fig:ACH_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{ACH_hop_acf.pdf}
  \caption{}\label{fig:ACH_hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_MET.pdf}
  \caption{}\label{fig:MET_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{MET_powerlaw.pdf}
  \caption{}\label{fig:MET_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{MET_hop_acf.pdf}
  \caption{}\label{fig:MET_hop_acf}
  \end{subfigure}
  \caption{Hop length distributions, dwell time distributions and hop autocorrelation functions
  respectively for urea (a-c), ethylene glycol (d-f), acetic acid (g-i), and methanol (j-l).
  Using urea as an example, (a) the distribution of hop lengths can be fit by a Gaussian in
  addition to a more general L\'evy stable distribution, though the L\'evy stable distribution
  does a better job of capturing the heavier tails and increased density near 0. We explore
  models fit to both distributions, as Gaussian hops are more convenient to model. (b) The 
  distribution of dwell times is fit well by a power law but it over-estimates the probability
  density at long dwell times. A power law truncated with an exponential cut-off better
  describes the probability of long dwell times in our simulations. (c) The hops are negatively
  correlated to their previous hop. In combination, (a) -- (c) support modeling solutes as either
  subordinated fractional Brownian or L\'evy motion.
  }\label{fig:anticorrelated_hops}
  \end{figure}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison.pdf}
  \caption{}\label{fig:gaussian_levy_comparison}
  \end{subfigure}
  % /figures/msddm_acf.py
  \begin{subfigure}{0.42\textwidth}
  \includegraphics[width=\textwidth]{msddm_acf.pdf}
  \caption{}\label{fig:msddm_acf}
  \end{subfigure}
  \caption{(a) The emission distributions of hop lengths are non-Gaussian and
	  heavy-tailed. Shown here is the emission distribution for transitions
	  between states. The maximum likelihood Gaussian fit severely
	  underestimates the empirical density of hops near and far from zero
	  while overestimating the density of hops at intermediate values. (b)
	  Jumps drawn from the transition distribution are negatively
	  correlated to each other. The normalized version of
	  Equation~\ref{M-eqn:flm_autocovariance} fits well to the data
	  suggesting FLM is an appropriate way to model jumps.}
	  \label{fig:msddm_emissions}
  \end{figure}

>>>>

  % mimic this
  \texttt{/setup/parameterize.py}            & 2.1 & Parameterize liquid crystal monomers and solutes with GAFF \\ \hline
  \texttt{/setup/build.py}                   & 2.1 & Build simulation unit cell \\ \hline
  \texttt{/setup/place\_solutes\_pores.py}   & 2.1 & Place equispaced solutes in the pore centers of a unit cell \\\hline
  \texttt{/setup/equil.py}                   & 2.1 & Equilibrate unit cell and run production simulation \\\hline
  \texttt{/analysis/solute\_partitioning.py} & 2.1 & Determine time evolution of partition of solutes between pores and tails \\\hline
  \texttt{/timeseries/msd.py}                & 2.2 & Calculate the mean squared displacement of solutes \\\hline
  \texttt{/analysis/sfbm\_parameters.py}     & 2.2 & Get subordinated fractional Brownian motion parameters by fitting to a solute's dwell and hop length distributions and positional autocorrelation function. \\\hline
  \texttt{/timeseries/ctrwsim.py}            & 2.2 & Generate realizations of a continuous time random walk with the user's choice of dwell and hop distributions \\\hline
  \texttt{/timeseries/forecast\_ctrw.py}     & 2.2 & Combines classes from \texttt{sfbm\_parameters.py} and \texttt{ctrwsim.py} to parameterize and predict MSD in one shot. \\\hline
  \texttt{/analysis/ markov\_state\_dependent\_dynamics.py} & 2.3 & Identify frame-by-frame state of each solute, construct a transition matrix and simulate realizations of the MSDDM model. \\\hline
  \texttt{/timeseries/mfpt\_pore.py}          & 2.4 & Simulate mean first passage time distributions using the AD approach. \\\hline 
  \end{tabular}

@@

  There are many methods for estimating the Hurst parameter for a time series.
  \cite{clegg_practical_2006} It can be a difficult task because 
  Equation~\ref{eqn:fbm_autocorrelation} decays slowly to zero, especially when 
  $H > 0.5$, meaning one needs to study large time lags with high frequency.

  Fortunately, from a mathematical perspective, all of our solutes show anti-correlated
  motion, so most of the information in Equation~\ref{eqn:fbm_autocorrelation} is contained
  within the first few time lags. In Figure~\ref{fig:hurst_autocorrelation}, we plotted
  Equation~\ref{eqn:fbm_autocorrelation} for different values of $H$. When $H > 0.5$, 
  Equation~\ref{eqn:fbm_autocorrelation} decays slowly to zero meaning one needs to study
  large time lags with high frequency in order to accurately estimate $H$ from the data. 
  However, when $H < 0.5$, the autocorrelation function quickly decays towards zero.

@@

  \textit{Determining where to truncate the hop distribution:} A pure
  L\'evy stable distribution has heavy tails which can lead to arbitrarily
  long hop lengths. Our distribution of hop lengths fits well to a L\'evy
  distribution near the mean, but under-samples the tails. In 
  Figure~\ref{fig:truncation_cutoff} we compare the empirically 
  measured transition emission distribution of the MSDDM for urea to its maximum likelihood fit to
  a L\'evy stable distribution. The ratio between the two distributions 
  at each bin is nearly 1 close to the center, indicating a near-perfect
  fit, larger than 1 slightly further from the center, suggesting that 
  we slightly over sample intermediate hop lengths, and below 1 far from 
  the center, indicating undersampling of extremely long hop lengths.
  Based on the plot, we chose a cut-off of 1 nm in order to compensate for
  over sampled intermediate hop lengths. We chose the same cut-off for all
  solutes.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{truncation_correction.pdf}
  \caption{}\label{fig:truncation_correction}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{hurst_correction.pdf}
  \caption{}\label{fig:hurst_correction}
  \end{subfigure}
  \caption{(a) We can accurately truncate the marginal distribution of FLM innovations by
  applying a correction to the input truncation parameter. We generated FLM sequences
  and truncated the initial L\'evy stable distribution (before Fourier transforming) at 
  a value of 0.5. After correlation structure is added, the width of the distribution 
  of fractional L\'evy noise decreases significantly. We corrected the input truncation
  parameter with our database resulting in a distribution close to the theoretical
  distribution (black dashed line) with a maximum value close to 0.5.
  (b) Correcting the Hurst parameter input to the algorithm of Stoev and Taqqu
  results in an FLM process with a more accurate correlation structure. We generated
  sequences with an input $H$ of 0.35. We estimated $H$ by fitting the autocorrelation
  function. Without the correction, $H$ is underestimated, meaning realizations are 
  more negatively correlated than they should be. }
  \end{figure}

@@

%  % /supporting_figures/demonstrate_hurst_correction.py
%  \begin{figure}
%  \centering
%  \includegraphics[width=0.5\textwidth]{hurst_correction.pdf}
%  \caption{Correcting the Hurst parameter input to the algorithm of Stoev and Taqqu
%  results in an FLM process with a more accurate correlation structure. We generated
%  sequences with an input $H$ of 0.35. We estimated $H$ by fitting the autocorrelation
%  function. Without the correction, $H$ is underestimated, meaning realizations are 
%  more negatively correlated than they should be.}\label{fig:hurst_correction}
%  \end{figure}

@@

  Figure~\ref{fig:anticorrelated_hops} demonstrates that all solutes exhibit the
  same kind of anti-correlated hopping and trapping behavior.
%  
  \begin{figure}[htb!]
  \centering
  % /figures/distributions_sfbm.py
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_URE.pdf}
  \caption{}\label{fig:URE_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{URE_powerlaw.pdf}
  \caption{}\label{fig:URE_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{URE_hop_acf.pdf}
  \caption{}\label{fig:URE_hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_GCL.pdf}
  \caption{}\label{fig:GCL_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{GCL_powerlaw.pdf}
  \caption{}\label{fig:GCL_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{GCL_hop_acf.pdf}
  \caption{}\label{fig:GCL_hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_ACH.pdf}
  \caption{}\label{fig:ACH_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{ACH_powerlaw.pdf}
  \caption{}\label{fig:ACH_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{ACH_hop_acf.pdf}
  \caption{}\label{fig:ACH_hop_acf}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison_anomalous_MET.pdf}
  \caption{}\label{fig:MET_hop_distribution_comparison}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{MET_powerlaw.pdf}
  \caption{}\label{fig:MET_powerlaw}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
  \includegraphics[width=\textwidth]{MET_hop_acf.pdf}
  \caption{}\label{fig:MET_hop_acf}
  \end{subfigure}
  \caption{Hop length distributions, dwell time distributions and hop autocorrelation functions
  respectively for urea (a-c), ethylene glycol (d-f), acetic acid (g-i), and methanol (j-l).
%  Using urea as an example, (a) the distribution of hop lengths can be fit by a Gaussian in
%  addition to a more general L\'evy stable distribution, though the L\'evy stable distribution
%  does a better job of capturing the heavier tails and increased density near 0. We explore
%  models fit to both distributions, as Gaussian hops are more convenient to model. (b) The 
%  distribution of dwell times is fit well by a power law but it over-estimates the probability
%  density at long dwell times. A power law truncated with an exponential cut-off better
%  describes the probability of long dwell times in our simulations. (c) The hops are negatively
%  correlated to their previous hop. In combination, (a) -- (c) support modeling solutes as either
%  subordinated fractional Brownian or L\'evy motion.
  }\label{fig:anticorrelated_hops}
  \end{figure}

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{gaussian_levy_comparison.pdf}
  \caption{}\label{fig:gaussian_levy_comparison}
  \end{subfigure}
  % /figures/msddm_acf.py
  \begin{subfigure}{0.42\textwidth}
  \includegraphics[width=\textwidth]{msddm_acf.pdf}
  \caption{}\label{fig:msddm_acf}
  \end{subfigure}
  \caption{(a) The emission distributions of hop lengths are non-Gaussian and
	  heavy-tailed. Shown here is the emission distribution for transitions
	  between states. The maximum likelihood Gaussian fit severely
	  underestimates the empirical density of hops near and far from zero
	  while overestimating the density of hops at intermediate values. (b)
	  Jumps drawn from the transition distribution are negatively
	  correlated to each other. The normalized version of
	  Equation~\ref{M-eqn:flm_autocovariance} of the main text fits well to the data
	  suggesting FLM is an appropriate way to model jumps.}
	  \label{fig:msddm_emissions}
  \end{figure}
